---
title: 'Lab 4 Deep Learning: iNaturalist'
author: "Halina Do-Linh"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(digest, 
                 tensorflow,
                 keras,
                 tidyverse)

# install Python into user space
(reticulate::miniconda_path()) # show the Python path
if (!file.exists(reticulate::miniconda_path()))
  reticulate::install_miniconda()

# install keras with tensorflow
if (!keras::is_keras_available())
  keras::install_keras()
```

# Set up

Here I am generating the 10 species I am going to use in my models, which are:

-   05992_Plantae_Tracheophyta_Liliopsida_Asparagales_Orchidaceae_Anacamptis_coriophora
-   06622_Plantae_Tracheophyta_Magnoliopsida_Asterales_Asteraceae_Arctotheca_prostrata
-   05437_Fungi_Ascomycota_Lecanoromycetes_Rhizocarpales_Rhizocarpaceae_Rhizocarpon_geographicum
-   00372_Animalia_Arthropoda_Insecta_Coleoptera_Passalidae_Odontotaenius_disjunctus
-   01212_Animalia_Arthropoda_Insecta_Lepidoptera_Geometridae_Euchlaena_serrata
-   07258_Plantae_Tracheophyta_Magnoliopsida_Brassicales_Brassicaceae_Lepidium_draba
-   06740_Plantae_Tracheophyta_Magnoliopsida_Asterales_Asteraceae_Delairea_odorata
-   09917_Plantae_Tracheophyta_Polypodiopsida_Polypodiales_Blechnaceae_Cranfillia_fluviatilis
-   02761_Animalia_Chordata_Actinopterygii_Cypriniformes_Catostomidae_Catostomus_commersonii
-   03624_Animalia_Chordata_Aves_Galliformes_Odontophoridae_Colinus_virginianus

```{r}
# path to folder containing species directories of images
dir_train_mini <- "/courses/EDS232/inaturalist-2021/train_mini"

# get list of directories, one per species (n = 10,000 species)
dirs_spp <- list.dirs(dir_train_mini, recursive = F)
n_spp <- length(dirs_spp)

# set seed (for reproducible results) 
# just before sampling (otherwise get different results)
# based on your username (unique amongst class)
Sys.info()[["user"]] %>% # sys.info pulls my user name
  digest::digest2int() %>% 
  set.seed()
i10 <- sample(1:n_spp, 10)

# show the 10 indices sampled of the 10,000 possible 
i10

# show the 10 species directory names
# saved as a variable here to make for loops below
species_10 <- basename(dirs_spp)[i10]

# show the 2 species directory names
i2 <- i10[1:2]
species_2 <- basename(dirs_spp)[i2]
```
Here I am creating base file paths to create my directories.

```{r}
# same file paths for both species 2 and species 10
original_dataset_dir <- "/courses/EDS232/inaturalist-2021/train_mini"

# base file paths
base_dir <- "/Users/halina/hd-eds-232/lab4-deepLearning"
```

Here I am creating my files paths and directories for my species.

```{r}
# create base file paths
train_dir_2 <- file.path(base_dir, "train_2")
train_dir_10 <- file.path(base_dir, "train_10")

validation_dir_2 <- file.path(base_dir, "validation_2")
validation_dir_10 <- file.path(base_dir, "validation_10")
  
test_dir_2 <- file.path(base_dir, "test_2")
test_dir_10 <- file.path(base_dir, "test_10")
```


```{r, eval=FALSE}
# create base directories (train, validate, test) for the two species and the ten species 
dir.create(train_dir_2)
dir.create(validation_dir_2)
dir.create(test_dir_2)
dir.create(train_dir_10)
dir.create(validation_dir_10)
dir.create(test_dir_10)

# create directories for two species
for (i in 1:length(species_2)){
  dir.create(file.path(train_dir_2, str_sub(species_2[[i]], start = 1, end = 5)))
  dir.create(file.path(validation_dir_2, str_sub(species_2[[i]], start = 1, end = 5)))
  dir.create(file.path(test_dir_2, str_sub(species_2[[i]], start = 1, end = 5)))
}
# create directories for all ten species
for (i in 1:length(species_10)){
  dir.create(file.path(train_dir_10, str_sub(species_10[[i]], start = 1, end = 5)))
  dir.create(file.path(validation_dir_10, str_sub(species_10[[i]], start = 1, end = 5)))
  dir.create(file.path(test_dir_10, str_sub(species_10[[i]], start = 1, end = 5)))
}
```

Here I am adding the images to my two species directories.

```{r, eval=FALSE}
# create test, validation, and training groups of images for two species
for(i in 1:length(species_2)){
  # create 5 groups of 10 random samples
  species_samples_2 <- replicate(5, sample(list.files(paste0(original_dataset_dir, "/", species_2[[i]]), full.names = TRUE), replace = FALSE, 10))
  ## train n = 30 ##
  train <- rbind(species_samples_2[,1], species_samples_2[,2], species_samples_2[,3])
  file.copy(from = train, 
            to = paste0(train_dir_2, "/", str_sub(species_2[[i]], start = 1, end = 5)))
  ## validation n = 10 ##
  validate <- species_samples_2[,4]
  file.copy(from = validate,
            to = paste0(validation_dir_2, "/", str_sub(species_2[[i]], start = 1, end = 5)))
  ## train n = 10 ##
  test <- species_samples_2[,5]
  file.copy(from = test,
            to = paste0(test_dir_2, "/", str_sub(species_2[[i]], start = 1, end = 5)))
}
```


Here I am adding the images to my ten species directories.

```{r, eval=FALSE}
# create test, validation, and training groups of images for ten species
for(i in 1:length(species_10)){
  species_samples_10 <- replicate(5, sample(list.files(paste0(original_dataset_dir, "/", species_10[[i]]), full.names = TRUE), replace = FALSE, 10))
  ## train n = 30 ##
  train <- rbind(species_samples_10[,1], species_samples_10[,2], species_samples_10[,3])
  file.copy(from = train, 
            to = paste0(train_dir_10, "/", str_sub(species_10[[i]], start = 1, end = 5)))
  ## validation n = 10 ##
  validate <- species_samples_10[,4]
  file.copy(from = validate,
            to = paste0(validation_dir_10, "/", str_sub(species_10[[i]], start = 1, end = 5)))
  ## train n = 10 ##
  test <- species_samples_10[,5]
  file.copy(from = test,
            to = paste0(test_dir_10, "/", str_sub(species_10[[i]], start = 1, end = 5)))
}
```

## Preprocessing the data for two species

```{r}
# All images will be rescaled by 1/255
test_datagen <- image_data_generator(rescale = 1/255)
train_datagen <- image_data_generator(rescale = 1/255)
validation_datagen <- image_data_generator(rescale = 1/255)

train_generator_2 <- flow_images_from_directory(
  # This is the target directory
  train_dir_2,
  # This is the data generator
  train_datagen,
  # All images will be resized to 150x150
  target_size = c(150, 150),
  batch_size = 5,
  # Since we use binary_crossentropy loss, we need binary labels
  class_mode = "binary") 

validation_generator_2 <- flow_images_from_directory(
  validation_dir_2,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "binary")
```

**Note: only 43 images out of 60....**

## Preprocessing the data for ten species

```{r}
train_generator_10 <- flow_images_from_directory(
  train_dir_10,
  train_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "categorical") 

validation_generator_10 <- flow_images_from_directory(
  validation_dir_10,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "binary")
```

**Note 116/100 and 266/300**

# 1. Two Species (binary classification) using neural net

## Building the network

```{r}
model_nn <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(150, 150)) %>%
  layer_flatten() %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units =  1, activation = "sigmoid")
```


```{r}
model_nn %>% compile(
  optimizer = "rmsprop",
  loss      = "binary_crossentropy",
  metrics   = c("accuracy"))
```


```{r}
history <- model_nn %>% fit(
    train_generator_2,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = validation_generator_2,
    validation_steps = 5)
```

```{r}
plot(history)
```

History results of neural net:

```{r}
history
```

## Evaluate Model

```{r}
test_generator <- flow_images_from_directory(
  test_dir_2,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
model %>% evaluate(test_generator, steps = 30)
```

# 2. Two Species (binary classification) - using convolutional neural net

## Building the network

```{r}
model_cnn <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dense(units = 512, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
```


```{r}
model_cnn %>% compile(
  optimizer = "rmsprop",
  loss      = "binary_crossentropy",
  metrics   = c("accuracy"))
```


```{r}
history_cnn <- model_cnn %>% fit(
    train_generator,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = validation_generator,
    validation_steps = 5)
```

```{r}
plot(history_cnn)
```

History results of convolutional neural net:

```{r}
history_cnn
```
## Evaluate Model

```{r}
test_generator <- flow_images_from_directory(
  test_dir_2,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "binary"
)
model_cnn %>% evaluate(test_generator, steps = 4) # 4 steps?? compare
```

# 3. Ten Species (multi-class classification) - neural net

## Building the network

```{r}
model_mc_nn <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(150, 150, 3)) %>%
  layer_flatten() %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units =  1, activation = "sigmoid")
```


```{r}
model_mc_nn %>% compile(
  optimizer = "rmsprop",
  loss      = "categorical_crossentropy",
  metrics   = c("accuracy"))
```


```{r}
history_mc_nn <- model_mc_nn %>% fit(
    train_generator,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = validation_generator,
    validation_steps = 5)
```

```{r}
plot(history_mc_nn)
```

## Evaluate Model

```{r}
test_generator <- flow_images_from_directory(
  test_dir_2,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "binary"
)
model_mc_nn %>% evaluate(test_generator, steps = 4) # 4 steps?? compare
```

# 4. Ten Species (multi-class classification) - convolutional neural net

```{r}
model_mc_nn <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dense(units = 512, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
```


```{r}
model_mc_nn %>% compile(
  optimizer = "rmsprop",
  loss      = "categorical_crossentropy",
  metrics   = c("accuracy"))
```


```{r}
history_mc_nn <- model_mc_nn %>% fit(
    train_generator,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = validation_generator,
    validation_steps = 5)
```

```{r}
plot(history_mc_nn)
```

## Evaluate Model

```{r}
test_generator <- flow_images_from_directory(
  test_dir_2,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 5,
  class_mode = "binary"
)
model_mc_nn %>% evaluate(test_generator, steps = 4) # 4 steps?? compare
```


