---
title: "Notes"
author: "Halina Do-Linh"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes



There is no relationship between these two metrics.
Loss can be seen as a distance between the true values of the problem and the values predicted by the model. Greater the loss is, more huge is the errors you made on the data.

Accuracy can be seen as the number of error you made on the data.

That means:

- a low accuracy and huge loss means you made huge errors on a lot of data
- a low accuracy but low loss means you made little errors on a lot of data
- a great accuracy with low loss means you made low errors on a few data (best case)

your situation: a great accuracy but a huge loss, means you made huge errors on a few data.


```{r, echo=F, eval=F}
# in Terminal:
#   cd /courses/EDS232; mkdir 'inaturalist-2021'
#   curl -o train_mini.tar.gz https://ml-inat-competition-datasets.s3.amazonaws.com/2021/train_mini.tar.gz
#   tar -xzf train_mini.tar.gz 

# librarian::shelf(
#   dplyr, glue, jsonlite, listviewer, purrr, readr, tidyjson, tidyr)
# 
# train_mini <- jsonlite::read_json("~/courses/train_mini.json")
# 
# write_meta <- function(m){
#   train_mini[[m]]  %>% 
#     tidyjson::spread_all() %>% 
#     tibble() %>% 
#     select(-document.id, -`..JSON`) %>% 
#     write_csv(
#       glue("~/Desktop/iNat/train_mini_{m}.csv"))
# }
# write_meta("images")
# write_meta("annotations")
# write_meta("categories")
```

Helpful `file.copy()` example

```{r}
current.folder <- "E:/ProjectDirectory/Data/"
new.folder <- "E:/ProjectDirectory/NewData/"
list.of.files <- list.files(current.folder, full.names = T)
# copy the files to the new folder
file.copy(list.of.files, new.folder, overwrite = T)

### my code ###
file.copy()
# current folder
original_orchid_dir
# orchid files
original_orchid_files <- list.files(original_orchid_dir, full.names = T)
# new folder
train_orchid_dir

# basenames is a good function to know
original_orchid_files <- list.files(original_orchid_dir, full.names = TRUE)
head(basename(original_orchid_files))
```


Model 

```{r}
# check if already fitted and saved model
if (!file.exists(mdl1_history_rds) | !file.exists(mdl1_h5)){
  # fit model
  history <- model %>% fit_generator(
    train_generator,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = validation_generator,
    validation_steps = 5)
  
  # save fitted model and fitting history
  # history %>% saveRDS(mdl1_history_rds)
  # model %>% save_model_hdf5(mdl1_h5)
} #else{
  # load previously fitted model
  # history <- readRDS(mdl1_history_rds)
  # model   <- load_model_hdf5(mdl1_h5)
#}
```

Here I am creating the file paths for the train, validatation, and test directories.

```{r}
original_dataset_dir <- "/courses/EDS232/inaturalist-2021/train_mini"

# base
base_dir <- "/Users/halina/hd-eds-232/lab4-deepLearning"
train_dir <- file.path(base_dir, "train")
validation_dir <- file.path(base_dir, "validation")
test_dir <- file.path(base_dir, "test")

# orchid 
train_orchid_dir <- file.path(train_dir, "orchid")
validation_orchid_dir <- file.path(validation_dir, "orchid")
test_orchid_dir <- file.path(test_dir, "orchid")

# aster
train_aster_dir <- file.path(train_dir, "aster")
validation_aster_dir <- file.path(validation_dir, "aster")
test_aster_dir <- file.path(test_dir, "aster")
```


Here I create the train, validation, and test directories for my current working directory, and my species orchid and aster.

```{r, echo=TRUE, results='hide', eval=TRUE}
# base
dir.create(train_dir)
dir.create(validation_dir)
dir.create(test_dir)

# orchid
dir.create(train_orchid_dir)
dir.create(validation_orchid_dir)
dir.create(test_orchid_dir)

# aster
dir.create(train_aster_dir)
dir.create(validation_aster_dir)
dir.create(test_aster_dir)
```



Here I added randomly sampled image files from `original_dataset_dir` to my train, validation, and test directories. I am splitting the original image files (50 total) for my species as follows:

- 30 images train
- 10 images validation
- 10 images test

```{r}
# directories for my species
original_orchid_dir <- "/courses/EDS232/inaturalist-2021/train_mini/05992_Plantae_Tracheophyta_Liliopsida_Asparagales_Orchidaceae_Anacamptis_coriophora"

original_aster_dir <- "/courses/EDS232/inaturalist-2021/train_mini/06622_Plantae_Tracheophyta_Magnoliopsida_Asterales_Asteraceae_Arctotheca_prostrata"

### orchid files ###
## train n = 30 ##
orchid_30 <- sample(list.files(original_orchid_dir, full.names = TRUE), 30) 
# randomly select 30 files using sample() and used list.files() to list the full names of the files from the file path I specified in original_orchid_dir
file.copy(from = orchid_30, to = train_orchid_dir)
## validation n = 10 ##
orchid_10_v <- sample(list.files(original_orchid_dir, full.names = TRUE), 10)
file.copy(from = orchid_10_v, to = validation_orchid_dir)
## test n = 10 ##
orchid_10_t <- sample(list.files(original_orchid_dir, full.names = TRUE), 10)
file.copy(from = orchid_10_t, to = test_orchid_dir)

### aster files ###
## train n = 30 ##
aster_30 <- sample(list.files(original_aster_dir, full.names = TRUE), 30) 
file.copy(from = aster_30, to = train_aster_dir)
## validation n = 10 ##
aster_10_v <- sample(list.files(original_aster_dir, full.names = TRUE), 10)
file.copy(from = aster_10_v, to = validation_aster_dir)
## test n = 10 ##
aster_10_t <- sample(list.files(original_aster_dir, full.names = TRUE), 10)
file.copy(from = aster_10_t, to = test_aster_dir)
```